AhirV2: from algorithms to hardware \\ A tutorial introduction

### 2011-05-25 Documents/AhirV2/AhirV2.tex
updated the documentation..

The options:
\begin{itemize}
\item {\bf -I n}: specifies that external references to memory
are to be mapped as if they are to an internal object whose size
is $n$ bytes.
\item {\bf -E obj-name} : specifies that the object to which
external references are mapped is to be named obj-name.
\end{itemize}
It is better if you use the {\bf -I} and {\bf -E} options.

@@

The options:
\begin{itemize}
\item {\bf -O} : if used, sequential statement blocks are parallelized
by doing dependency analysis.
\item {\bf -C} : if used, a C stub is created for every module that
is not called from within the system.  These stubs can be used to
interface to a VHDL simuator (or even drive hardware) to verify
the VHDL code generated by downstream tools.
\item {\bf -I obj-name} : if specified, all external memory references
are considered as being directed at the storage object named obj-name.
\end{itemize}

@@

Takes a collection of \vC descriptions and converts them to
VHDL.
\begin{verbatim}
vc2vhdl [-O] [-C] -t foo [-t bar -t bar2 ...]\
         -f file1.vc -f file2.vc ...  > system.vhdl
\end{verbatim}

The options:
\begin{itemize}
\item {\bf -t} : to specify the modules which are to be 
accessible from the ports of the generated VHDL system.
Multiple top-level modules can be specified in this way.
\item {\bf -f} : specifies the \vC files to be analyzed. 
Multiple \vC files may be specified.
\item {\bf -O} : optimize the generated VHDL by compacting
the control-path.  This does not change the resulting
hardware, but makes the generated VHDL file smaller.
\item {\bf -C} : the VHDL code has a system test bench which
interfaces to foreign code using a VHPI/Modelsim-FLI interface.
If this is not specified, the  generated test bench simply
instantiates the system and starts all top-level modules
off (you will need to fill in your own test bench here).
The C testbench is usually easier to write (it probably
already exists in the form of the original program).
\end{itemize}

@@

Finally, we take the \vC description and convert it to
VHDL
\begin{verbatim}
vc2vhdl -t add -f add.o.aa.vc | vhdlFormat > system.vhdl
\end{verbatim}
This produces a VHDL implementation of the system with
{\bf add} marked as a top-level module.  The VHDL that
is produced is too voluminous to reproduce here, but
the top-level system entity is
\begin{verbatim}
entity test_system is  -- system
  port (--
    add_a : in  std_logic_vector(31 downto 0);
    add_b : in  std_logic_vector(31 downto 0);
    add_ret_val_x_x : out  std_logic_vector(31 downto 0);
    add_tag_in: in std_logic_vector(0 downto 0);
    add_tag_out: out std_logic_vector(0 downto 0);
    add_start : in std_logic;
    add_fin   : out std_logic;
    clk : in std_logic;
    reset : in std_logic); --
  --
end entity;
\end{verbatim}
There are ports corresponding to the arguments of the top-level module,
and the add\_start/add\_fin is a handshake pair.  One sets of the inputs,
starts the system and waits until the fin is asserted.  After the fin
is asserted, one has the return value of at the appropriate port.

@@

Both options are supported in the AhirV2 flow through
the utility {\bf AaLinkExtMem} described earlier.

@@

For this example, if you want to keep the external memory
outside, you will have to go through the following sequence
\begin{verbatim}
# use clang (or llvm-gcc) to generate llvm-byte-code
clang -std=gnu89 -emit-llvm -c foo.c
#
# disassemble so that you can make sense of the llvm bc.
llvm-dis foo.o
#
# OK, now take the llvm byte code
# and generate an Aa description.
# use the storageinit option to initialize
# global storage.
# (the pipe to vcFormat is to prettify the output)
 llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
#
#
# Do an Aa -> Aa transformation: map external
# memory to a storage area inside the system...
# -I 1024 says that the amount of memory that will be
# referred to is 1024 bytes.
# -E mempool says that the storage object corresponding
# to external memory is named mempool.
AaLinkExtMem  -I 1024 -E mempool foo.o.aa | vcFormat\
               > foo.o.memlinked.ExternalInside.aa
#
# Now take the Aa code and generate a virtual
# circuit..
# the -O flag does dependency analysis in straight-line
# code and parallelizes it.
# the -I mempool option says that external memory is
# to be mapped inside the system to object mempool..
#
Aa2VC -O -I mempool foo.o.memlinked.ExternalInside.aa\
           | vcFormat > foo.o.memlinked.ExternalInside.aa.vc
#
# finally, generate vhdl from the vc description.
# note that you will have to mark mem_load__ and mem_store__
# as top-level modules, so that the external world can
# access its memory pool inside the system.
#
vc2vhdl -O -t foo -t mem_load__ -t mem_store__ \ 
     -f foo.o.memlinked.ExternalInside.aa.vc\
       | vhdlFormat > foo_o_aa_memlinked_external_inside_vc.vhdl
\end{verbatim}

@@

The examples folder contains some explanations which might
make things easier to understand.

>>>>

The options:
\begin{itemize}
\item {\bf -I n}: specifies that external references to memory
are to be mapped as if they are to an internal object whose size
is $n$ bytes.
\item {\bf -E obj-name} : specifies that the object to which
external references are mapped is to be named obj-name.
\end{itemize}
It is recommended that you use the {\bf -I} and {\bf -E} options.

@@

The options:
\begin{itemize}
\item {\bf -O} : if used, sequential statement blocks are parallelized
by doing dependency analysis.
\item {\bf -C} : if used, a C stub is created for every module that
is not called from within the system.  These stubs can be used to
interface to a VHDL simuator (or even drive hardware) to verify
the VHDL code generated by downstream tools.
\item {\bf -I obj-name} : if specified, all external memory references
are considered as being directed at the storage object named obj-name.
If not specified, then the tool will throw an error if it finds
a pointer dereference that cannot be resolved as pointing only to
storage objects declared inside the \Aa program.
\end{itemize}

@@

Takes a collection of \vC descriptions and converts them to
an AHIR system described in VHDL.
\begin{verbatim}
vc2vhdl [-O] [-C] -t/-T foo [-t/-T bar -t/-T bar2 ...]\
         -f file1.vc -f file2.vc ...  > system.vhdl
\end{verbatim}

The options:
\begin{itemize}
\item {\bf -t} : to specify the modules which are to be 
accessible from the ports of the generated VHDL system.
Such modules have to be top-level (that is, they cannot
be called from within the program).
Multiple top-level modules can be specified in this way.
The control and argument ports for these modules are
visible at the interface of the generated AHIR system.
\item {\bf -T} : to specify top-level modules which are to be 
free-running inside the AHIR system.
Multiple top-level modules can be specified in this way.
Such modules do not have any arguments and do not return
any values.  Their only mechanism of communication with
the world outside the AHIR system is through pipes.
The control ports for these modules are
{\bf not visible} at the interface of the generated AHIR system.
In the AHIR system, these modules are started on reset
and are run forever (restarted after they finish, forever).
\item {\bf -f} : specifies the \vC files to be analyzed. 
Multiple \vC files may be specified.  An object must be
defined before it is used, so the \vC files must be 
specified in the correct order.
\item {\bf -O} : optimize the generated VHDL by compacting
the control-path.  This does not change the resulting
hardware, but makes the generated VHDL file smaller.
\item {\bf -C} : the VHDL code has a system test bench which
interfaces to foreign code using a VHPI/Modelsim-FLI interface.
If this is not specified, the  generated test bench simply
instantiates the system and starts all top-level modules
off (you will need to fill in your own test bench here).
The C testbench is usually easier to write (it probably
already exists in the form of the original program).
\item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
with the -s option, then the generated testbench (if -C is specified)
uses the VHPI interface to link with foreign code.  Otherwise,
the generated testbench (if -C is specified) uses the Modelsim FLI
interface to link with foreign code.
\end{itemize}

@@

Finally, we take the \vC description and convert it to
VHDL
\begin{verbatim}
vc2vhdl -t add -f add.o.aa.vc | vhdlFormat > system.vhdl
\end{verbatim}
This produces a VHDL implementation of the system with
{\bf add} marked as a top-level module.  The VHDL that
is produced is too voluminous to reproduce here, but
the top-level system entity is
\begin{verbatim}
entity test_system is  -- system
  port (--
    add_a : in  std_logic_vector(31 downto 0);
    add_b : in  std_logic_vector(31 downto 0);
    add_ret_val_x_x : out  std_logic_vector(31 downto 0);
    add_tag_in: in std_logic_vector(0 downto 0);
    add_tag_out: out std_logic_vector(0 downto 0);
    add_start : in std_logic;
    add_fin   : out std_logic;
    clk : in std_logic;
    reset : in std_logic); --
  --
end entity;
\end{verbatim}
There are ports corresponding to the arguments of the top-level module,
and the add\_start/add\_fin is a handshake pair.  One sets of the inputs,
starts the system and waits until the fin is asserted.  After the fin
is asserted, one has the return value of {\bf add} at the appropriate port.

@@

Both options are supported in the AhirV2 flow through
{\bf AaLinkExtMem}.

@@

For this example, if you want to keep the external memory
inside, you will have to go through the following sequence
\begin{verbatim}
# use clang (or llvm-gcc) to generate llvm-byte-code
clang -std=gnu89 -emit-llvm -c foo.c
#
# disassemble so that you can make sense of the llvm bc.
llvm-dis foo.o
#
# OK, now take the llvm byte code
# and generate an Aa description.
# use the storageinit option to initialize
# global storage.
# (the pipe to vcFormat is to prettify the output)
 llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
#
#
# Do an Aa -> Aa transformation: map external
# memory to a storage area inside the system...
# -I 1024 says that the amount of memory that will be
# referred to is 1024 bytes.
# -E mempool says that the storage object corresponding
# to external memory is named mempool.
AaLinkExtMem  -I 1024 -E mempool foo.o.aa | vcFormat\
               > foo.o.memlinked.ExternalInside.aa
#
# Now take the Aa code and generate a virtual
# circuit..
# the -O flag does dependency analysis in straight-line
# code and parallelizes it.
# the -I mempool option says that external memory is
# to be mapped inside the system to object mempool..
#
Aa2VC -O -I mempool foo.o.memlinked.ExternalInside.aa\
           | vcFormat > foo.o.memlinked.ExternalInside.aa.vc
#
# finally, generate vhdl from the vc description.
# note that you will have to mark mem_load__ and mem_store__
# as top-level modules, so that the external world can
# access its memory pool inside the system.
#
vc2vhdl -O -t foo -t mem_load__ -t mem_store__ \ 
     -f foo.o.memlinked.ExternalInside.aa.vc\
       | vhdlFormat > foo_o_aa_memlinked_external_inside_vc.vhdl
\end{verbatim}

@@
